The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates

### 2025-02-15 PediatricResearchSubmission_Oct2024/ReviewerResponse_2024Dec03/ResponsetoReviewers.md

---
title: Response to Reviewers
geometry: margin=1in
---

Reviewer's original suggestions are in quote blocks and in **bold**  
Our response follows in normal font, and changes to the document are in quote blocks and normal font.

# Editor Comments to Author

## Section Editor: 1

> **Comments to the Author:**  
> **This manuscript by Weber et al. is focused on oxygenation quantification in the SSS and CCV in 19 premature infants utilizing susceptibility mapping and paramagnetic separation, which is compared directly and thoroughly with the preceding literature using a variety of techniques. This is significant in that it may provide an additional, and hopefully more accurate, way to measure oxygenation in premature infants. Overall this is an impressive and well planned study.**

Thank you for your kind words and for taking the time to review our manuscript.

---

> **The main issue which needs to be addressed revolves around the conclusions, and that the presented data do not support that this has superior accuracy. It is important because some of the findings contradict previous studies. The language here needs to be softened**

We agree. It is important that our conclusions do not overstate the evidence provided. Based on the editor's recommendation, we have now softened our language in the abstract, impact statement, and conclusion:

Abstract:

> While paramagnetic component decomposition yielded SSS values closer to those found in the literature, it increased variability. No significant oxygenation differences were found between the SSS and CCV, contrasting with prior studies.

Impact Statement:

> This study evaluated the use of QSM and its paramagnetic components to measure cerebral oxygenation in neonates.

Conclusion:

> This study aimed to evaluate how the use of susceptibility separation on preterm neonatal QSM images could be used in determining the oxygenation of cerebral venous vessels

---

> **In the methods, there are 3 lines where "\*\*\*" appears, I am not sure what that is supposed to represent (lines 102, 104, 117).**

These asterisks' were included in order to abide by what we believed *Pediatric Research's* submission guidelines were, to anonymize the paper from reviewers. These will be replaced with the identifying information once accepted. Or, if we made a mistake, we are happy to de-anonymize them for the reviewers.

---

> **Requires additional clarification related to imaging analysis and how it leads to their conclusion**

We agree with the editor and with the reviewers below who made similar comments. Please see our responses below for a detailed summary of the changes we made.

# Reviewer(s)' Comments to Author

## Reviewer: 1

> **Comments to the Author**  
> **Summary:**  
> **This study aims to quantify oxygenation in the superior sagittal sinus and in central cerebral veins in a cohort of 19 preterm neonates using quantitative susceptibility mapping and the separation of its paramagnetic components.**  
> **The authors observed no significant differences between the oxygenation of the SSS and CCV, which contradicts previous results reported in the literature. They reported that the accuracy of the paramagnetic source separation was higher than QSM, but at the cost of an increased variability.**  
> **The methods of the manuscript are well detailed, and the results are appropriately compared to the literature. A few aspects require some clarification, see below.**

We thank the reviewer for taking the time to read our manuscript and for making suggestions to help improve the final result.

---

> **One limitation of this work is the lack of sufficient data to support the claim of increased accuracy with susceptibility sources separation compared to QSM. In the absence of a ground truth and with results that contradict literature, this work does not support the increased accuracy claimed. The conclusion of the work should be re-written to better reflect the actual results.**

We agree with the reviewer. Without a ground truth to compare to, it is difficult to say if our results reflect a more accurate measure. We have now included wording saying as much, and rewritten the conclusion as follows:

> This study aimed to evaluate how the use of susceptibility separation on preterm neonatal QSM images could be used in determining the oxygenation of cerebral venous vessels

The impact statement was also changed:

> This study evaluated the use of QSM and its paramagnetic components to measure cerebral oxygenation in neonates.

Along with the abstract:

> While paramagnetic component decomposition yielded SSS values closer to those found in the literature, it increased variability. No significant oxygenation differences were found between the SSS and CCV, contrasting with prior studies.

---

> **Specific Issues:**  
> **Abstract:**  
> **1. Please re-write the first sentence of the conclusion section of the abstract to better reflect the results presented in this work, which do not fully support the current claim of increased accuracy with susceptibility source separation.**

The first sentence of the conclusion of the abstract has been changed to better reflect our findings:

> While paramagnetic component decomposition yielded SSS values closer to those found in the literature, it increased variability. No significant oxygenation differences were found between the SSS and CCV, contrasting with prior studies.

---

> **Introduction:**  
> **2. It is mentioned that methods to assess whole brain oxygenation have been developed for adults and that they are starting to be explored in neonates. Please detail what are the challenges for transferring these methods to neonate population.**

We have adjusted this section to now read:

> For the preceding reasons, non-invasive MRI-based techniques are actively being explored to assess regional and whole-brain blood oxygenation. While MRI-based methods have been developed for adults (9-11), their application in neonates is only beginning to be explored (12-16). This delay is likely due to the unique challenges posed by neonates, including their smaller anatomies, distinct hemodynamic profiles, susceptibility to motion artifacts, and the difficulties associated with recruiting this population for research.

---

> **Method:**  
> **Image acquisition**  
>**3. Please state in the text the plane of acquisition for each sequence.**

We now include the plane of acquisition for each sequence in the text of the Image Acquisition section.

---

>**4. Table 1: The phase encoding row (2nd row) does not provide the phase encoding information. This row should be split in two, one for the acquisition plane information (axial, coronal, sagittal), and one for the actual phase encoding direction in plane (ex: right-left, antero-posterior, or head-foot).**

We thank the reviewer for catching this. We have now updated the table to reflect these changes by adding a row for Acquisition plane, and fixing the Phase-encoding direction

---

> **Image analysis:**  
> **5. Briefly explain why the fifth echo was selected to generate the preliminary brain mask.**

The last (fifth) echo was used to create the brain mask as the last echo will have the greatest signal decay from air-bone-tissue interfaces. This is important to use in order to properly avoid introducing aliasing in the inverse problem. We now include an explanation of this in our methods for readers:

> The last echo is used to generate the brain mask as it reliably removes artifacts from air-tissue and bone-tissue interfaces, e.g., sinuses, without the need for manual erosion. The reason for this is at longer echo times, tissues with rapid signal decay (such as bone, air, and sinuses) lose their MRI signal due to dephasing caused by greater magnetic field inhomogeneities.

---

> **6. Please clarify if all echoes were used for the preliminary QSM calculation.**

We apologise for the confusing way our methods section was written. The methods now describe that the fifth echo was used for mask creation; the STI suite used all five echoes to create five separate QSM images; the 3rd, 4th and 5th QSM maps were then averaged to create the final QSM image for which chi values are taken. The $\chi$-separation toolbox outputs a single map for diamagnetic, paramagnetic, and QSM. Please see the tracked changes in the methods section, and our response to #8 below.

---

> **7. Can the authors expand on the motivation for choosing the STI software? What does ‘cleanest images’ refer to exactly (accuracy, artifact free, etc.)? Please also add a reference to support this statement.**

The STI software appeared to give the best images, based on a visual check by the authors. As this was based on our opinion, we cannot provide a reference for this.  
We now include an explanation of this in the methods:

> STI Suite (v. 3.0), was used to process the final QSM images as it produced the images with the least amount of artifacts (based on a visual assessment by the authors) without eroding the SSS.

---

> **8. Can the authors explain why only the last three echoes were used for the STI QSM calculation?**

This was an error in reporting we have now fixed. Please see the response below.

---

> **9. The methods description mentions that the last three echoes only were used for the STI calculation, but then in the parameters listed in the text, it is mentioned that a TE1=5 ms was used, which is the first TE of the sequence. Can the authors clarify this?**

We apologise for this. The way it was written was not clear. We used all five echoes to produce a five echo QSM image, but then averaged the last three echoes to produce the final QSM image. The section now reads:

> The finalized brain mask along with the magnitude and phase images were used in STI Suite along with the following parameters: 0.9766 x 0.9766 x 1 mm3 voxel size, 5 ms TE1, 5.24 ms TE, and 77.4 ms sum TE, B0 strength = 3, and B0 direction = [0, 0, 1]. The 3D GRE data option was selected for the phase processing stage, and STAR-QSM was selected for the QSM stage. The last three echoes of the QSM were then averaged to create the final QSM image using fslmaths, as the accumulation of phase due to susceptibility is small in early echoes and artifacts dominate the phase.

---

> **10. In the paragraph describing the methods for the paramagnetic source separation, please list the TEs in ms unit instead of s for consistency with the rest of the manuscript. Also in the same paragraph, please use the acronym introduce earlier in the text for delta TE for consistency.**

We thank the reviewer for catching these. We originally stated the TEs in seconds in this section as that is what the paramagnetic source software uses when entering units. However, we believe that using consistent units throughout the manuscript would be best. We have made this correction in this section.

---

> **11. Can the authors include an illustration of the final SSS and CCV masks used to calculate mean susceptibility in these regions?**

Yes of course. We agree that this would be a good addition to the paper.

Please see the new figure now numbered Figure 2.

---

> **Results:**  
> **12. Figure 2: Please add a label to the colorbars (ex: chi (ppm)). Also, please clarify in the caption what the x/y/z label on each figure refers to.**

"$\chi$ (ppm)" has now been added to the two colorbars. In addition, we now include the following in the caption:

> The color bars in (b) and \(c\) indicates the range of susceptibility $\chi$ values in parts per million.

The caption also now includes a sentence at the end that reads:

> y, x and z values represent the slice number in each plane (coronal, sagittal, and axial, respectively).

---

> **13. Figure 3: In the text referring to fig 3, the graphs presented are referred to as boxplots, versus in the figure caption, they are referred to as violin plots. Please make sure to consistently use the same term.**

We apologise for missing this. The text has been changed to reflect the appropriate plot name: 'violin plots'.

> Violin plots of the comparisons are shown in Figure 3.

---

> **Discussion**  
> **14. The authors mention being the first to explore the susceptibility source separation technique in neonates, while it has been done in adult population. Please discuss the challenges of transferring this technique to a neonate population and how these were address in this work.**

Often times new techniques will first be used in adults before eventually being used in neonatal populations. Part of this is due to the relative difficulty in acquiring ethics and recruiting subjects for neonatal research. My lab has a focus on neonatal brain development and I happen to be interested in QSM. We believe we are the first to perform this type of analysis in a neonatal population simply due to a combination of luck, access to data, and timing. However, please see our response to #2 above where we now include some of the challenges in scanning neonates in the introduction.

---

> **15. Can the authors discuss why, based on their results, QSM underestimates the susceptibility compared to the paramagnetic component separation technique?**

Reading over the manuscript, we realise this is a very important concept that should have been included in the introduction. We now include the following sentence near the end of the introduction:

> Furthermore, QSM tends to underestimates parametric components due to the inclusion of diamagnetic tissue, and vice versa, as the opposing magnetic susceptibilities effectively subtract from one another.

### 2025-02-15 index.qmd
yaml adjustments
final

---
title: "The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates"
titlerunning: "SVO2"
authors:
  - name: Thomas Gavin Carmichael
    orcid: 0009-0008-6849-5333
    corresponding: false
    email: tgcarmichael@outlook.com
    roles:
      - writing - original draft
      - formal analysis
      - methodology
      - validation
      - visualization
    affiliations:
      - ref: 1
      - ref: 2
    degrees:
      - HBSc
  - name: Alexander Rauscher
    orcid: 0000-0002-1961-8252
    email: rauscher@physics.ubc.ca
    corresponding: false
    roles:
      - writing - review & editing
    affiliations:
      - ref: 3
    degrees:
      - PhD
      - MSc
  - name: Ruth E Grunau
    orcid: 0000-0002-5428-9212
    corresponding: false
    email: rgrunau@mail.ubc.ca
    roles:
      - writing - review & editing
      - funding acquisition
    affiliations:
      - ref: 2
      - ref: 3
  - name: Alexander Mark Weber
    orcid: 0000-0001-7295-0775
    corresponding: true
    email: aweber@bcchr.ca
    roles:
      - project administration
      - supervision
      - validation
      - visualization
      - resources
      - methodology
      - formal analysis
      - funding acquisition
      - writing - review & editing
      - conceptualization
      - data curation
      - investigation
    affiliations:
      - ref: 2
      - ref: 3
    degrees:
      - PhD
      - MSc
affiliations:
  - id: 1
    name: Integrated Sciences, The University of British Columbia, Vancouver, BC, Canada
  - id: 2
    name: BC Children's Hospital Research Institute, The University of British Columbia, Vancouver, BC, Canada
  - id: 3
    name: Pediatrics, The University of British Columbia, Vancouver, BC, Canada

@@

abstract: |
  **Background**: Quantitative susceptibility mapping (QSM), an magnetic resonance imaging (MRI) modality sensitive to deoxyhemoglobin, is a promising method for measuring cerebral oxygenation in human neonates. Paramagnetic sources, like deoxyhemoglobin, however, can be obscured by diamagnetic sources such as water and myelin. This study evaluated whether QSM images, or isolated paramagnetic components, are more accurate for measuring oxygenation of cerebral veins of preterm neonates, and explored oxygenation differences between the major cerebral veins.

@@

  **Conclusion**: While paramagnetic component decomposition yielded more accurate SSS values, it increased variability. No significant oxygenation differences were found between the SSS and CCV, contrasting with prior studies.

bibliography: Gavin_Thesis_Ref.bib
csl: pediatric-research.csl

- This study evaluated the accuracy of using QSM and its paramagnetic components to measure cerebral oxygenation in neonates.
- By comparing susceptibility-derived oxygen saturation (SvO~2~) in the superior sagittal sinus (SSS) and central cerebral veins (CCV), it adds to the field of neonatal cerebral oxygenation measurement.
- Decomposing QSM into paramagnetic components shows potential for improving SvO~2~ accuracy, particularly in the SSS, though variability remains a challenge.
- The results suggest no significant oxygenation difference between the SSS and CCV, contrasting with previous findings, indicating a need for further research on neonatal venous oxygenation.

@@

Authors AMW and REG were co-primary applicants for a BC Children’s Hospital Research Institute - Brain, Behaviour and Development Catalyst Grant ($20,000). AMW was supported by an establishment award from BCCHRI. Scanning was partly funded through a special award to AMW from BCCHRI .

@@

For the preceding reasons, non-invasive MRI-based techniques are actively being explored to assess regional and whole-brain blood oxygenation. MRI based methods have been developed for adults [@jainRapidMagneticResonance2011; @luQuantitativeEvaluationOxygenation2008; @xuNoninvasiveQuantificationWholebrain2009] and are only beginning to be explored in neonates [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jainCerebralOxygenMetabolism2014; @jiangVesselspecificQuantificationNeonatal2019]. These methods almost all rely on T2 relaxation to estimate CSvO~2~ [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jiangVesselspecificQuantificationNeonatal2019] with the exception of @jainCerebralOxygenMetabolism2014 which used susceptometry [@jainMRIEstimationGlobal2010]. One limitation of these T2 relaxation methods, however, is the fact that SvO~2~ is often measured using a single imaging slice, averaging values across several voxels, and only in the superior sagittal sinus (SSS). In the case of @jainCerebralOxygenMetabolism2014, they obtained regional and whole-brain data, but with thick slices (5mm), and still only estimated CSvO~2~ in the SSS. An alternative MRI method using quantitative susceptibility mapping (QSM) has been proposed, which can measure SvO~2~ regionally and across the whole-brain at high resolution (< 1mm^3^ per voxel) [@weberQuantitativeSusceptibilityMapping2021]. However, this method left room for improvement, as it removed the SSS (averaging CSvO~2~ across the internal veins), and required an arbitrary threshold value of 0.15 ppm in order to acquire realistic results [@weberQuantitativeSusceptibilityMapping2021].

@@

Participant data comes from a previous study *** <!-- [@zhu-etal-cmro2] (submitted) -->. Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at *** <!-- BC Women’s Hospital -->. Recruitment took place over a span of one year, from February 2021 to January 2022, facilitated by a dedicated research nurse. Parents of eligible infants were approached by the research nurse prior to discharge from the NICU to explain the study objectives and seek their consent for participation. Infants meeting the criteria for inclusion were scanned for the study if they had already been discharged from the NICU, were in stable condition, and had reached a term equivalent age of 37 to 44 weeks GA. However, certain exclusion criteria were applied to ensure the homogeneity and integrity of the study sample: infants were excluded if there was clinical evidence of a congenital malformation or syndrome, a TORCH infection, or ultrasound evidence of large parenchymal hemorrhagic infarction (\>2 cm, Grade 4 intraventricular hemorrhage).

@@

MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750 scanner (scanner software version DV26.0_R03) equipped with a SREE Medical Systems (Cleveland, OH, USA) single-channel neonatal head coil (@tbl-mri). The scans were conducted at the *** <!-- BC Children -->’s MRI Research Facility. Prior to the scanning procedure, subjects were carefully prepared by a research nurse. Swaddling and feeding were used to ensure the comfort and cooperation of the subjects during the scan. Importantly, no sedatives or invasive markers were utilized throughout the procedure. Subjects were placed within a specially designed SREE Medical Systems MRI compatible incubator, which facilitated both safety and motion minimization. Molded foam was strategically positioned around the head and body within the incubator to further restrict subject movement. To protect against potential hearing damage, ear plugs were employed during the scanning process. Additionally, a pulse oximeter was affixed to the subject’s foot to monitor arterial oxygen saturation and heart rate throughout the scan.

```{r}
#| label: tbl-mri
#| tbl-cap: "**Technical parameters for MR imaging pulse sequences**"
df <- data.frame(Scan = character(), T1w = character(), T2w = character(), pcASL = character(), SWI = character(), stringsAsFactors = FALSE)
df[1,] <- c("Sequence", "3D FSPGR", "3D CUBE", "Multi-shot 3D fast spin-echo", "3D spoiled GRE flow-compensated")
df[2,] <- c("Phase-encoding direction", "Coronal", "Sagittal", "Axial", "Axial")
df[3,] <- c("TR (ms)", "7.74", "2,300", "4,680", "30.9")
df[4,] <- c("TE (ms)", "2.97", "66.29", "10.55", "5 echoes; first echo: 5; echo spacing: 5.24")
df[5,] <- c("Flip angle", "12\U00B0", "90\U00B0", "111\U00B0", "20\U00B0")
df[6,] <- c("FOV (cm)", "20", "20", "24", "25")
df[7,] <- c("Acquisition matrix", "512 x 512", "256 x 256", "128 x 128", "256 x 256")
df[8,] <- c("In-plane resolution (mm)", "0.39 x 0.39", "0.78 x 0.78", "1.875 x 1.875", "0.977 x 0.977")
df[9,] <- c("Slice thickness (mm)", "1", "1", "4", "2, reconstructed to 1 with zero filling (ZIP2)")
df[10,] <- c("Number of slices", "126", "106", "50", "92")
df[11,] <- c("Additional parameters", "n/a", "n/a", "1,450 ms label period;\n 2,025 ms pulse label;\n 24 control-label pairs", "n/a")
df[12,] <- c("Scan duration", "4 min 39 s", "5 min 1 s", "5 min 26 s", "5 min 29 s")

@@

if (knitr::is_latex_output()) {
colnames(df)[1] <- ""
df[11,] <- linebreak(df[11,])
df |>
  kbl(format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    align = "l",
    escape = FALSE) |>
    kable_styling(font_size = 8, position = "center", latex_options = c("hold_position", "scale_down"), full_width = FALSE) |>
    footnote(general_title = "", 
           footnote_as_chunk = TRUE,
           threeparttable = TRUE,
           general = footnotetext) |>
    column_spec(1, width="8em") |>
    column_spec(4, width="9em") |>
    column_spec(5, width="9em")
} else {
df |> 
    mutate(across(everything(), ~ str_replace_all(., "\n", "<br>"))) |> 
    gt(rowname_col = "Scan") |> tab_footnote(footnotetext) |>
    fmt_markdown(columns = TRUE) |>
    tab_options(quarto.disable_processing = TRUE)
}
```

The MRI scan protocol comprised of the following sequences: a T1-weighted scan, a T2-weighted scan, a pseudo-continuous arterial spin labeling (ASL) scan [@alsopRecommendedImplementationArterial2015], a multi-echo susceptibility-weighted imaging scan [@denkSusceptibilityWeightedImaging2010], and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence. The DWI sequence was not used for the present study.

@@

The raw DICOM files acquired from the scanning procedure were converted to NIfTI (Neuroimaging Informatics Technology Initiative) format using Chris Rorden’s `dcmniix` tool [@liFirstStepNeuroimaging2016]. SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past *** <!-- [@weberQuantitativeSusceptibilityMapping2021] -->. A step-by-step summary of the pipeline used is shown in @fig-graph.

First, the fifth echo (TE = 25.96 ms) SWI magnitude file was processed using FSL’s (v. 6.0.7.3) [@woolrichBayesianAnalysisNeuroimaging2009] `fslroi`, `fslmaths`, and `bet` [@smithFastRobustAutomated2002] to create a preliminary brain mask, similar to our previous efforts, which does not contain the SSS. `Fslroi` was used to isolate the fifth echo of the magnitude data, which was then squared using `fslmaths` and the option `-sqr`. Squaring the magnitude image was found to dramatically improve subsequent brain extraction. The resulting image was then used to create the preliminary brain mask using bet with the options `-m` and `-R`. The former flag generated a binary brain mask, while the latter performed a more robust brain centre estimation. The brain mask was then dilated by 7 voxels using `fslmaths` and the options `-kernel boxv` and `-dilM` in order for the dilated mask to contain the SSS (along with unwanted tissue as well). This mask was then used, along with the phase images, in a MATLAB script for QSM calculation from Christian Kames [@kamesRapidTwostepDipole2018] to produce a preliminary QSM image that contained the SSS, albeit with fairly low signal-to-noise ratio and other unwanted tissue. Given the high contrast in voxel intensity between the SSS and surrounding tissue, the select by intensity tool in `FSLeyes` [@mccarthyFSLeyes2023] was then used to segment the SSS from the QSM image and create a 3D mask of the selected region. Using `fslmaths` and the options `-add` and `-bin`, the SSS mask was then combined with the original brain mask of the fifth echo. This resulted in a brain mask that contained only brain and SSS signal. Finally, this mask was used in a final QSM post-processing step to create a QSM image that includes the SSS while maintaining a high signal-to-noise ratio, making it suitable to obtain accurate susceptibility values.

>>>>

---
title: "The Application of Magnetic Susceptibility Separation for Measuring Cerebral Oxygenation in Preterm Neonates"
titlerunning: "SVO2"
authors:
  - name: Thomas Gavin Carmichael
    orcid: 0009-0008-6849-5333
    corresponding: false
    email: tgcarmichael@outlook.com
    roles:
      - writing - original draft
      - formal analysis
      - methodology
      - validation
      - visualization
    affiliations:
      - ref: 1
      - ref: 2
    degrees:
      - HBSc
  - name: Alexander Rauscher
    orcid: 0000-0002-1961-8252
    email: rauscher@physics.ubc.ca
    corresponding: false
    roles:
      - writing - review & editing
    affiliations:
      - ref: 3
    degrees:
      - PhD
      - MSc
  - name: Ruth E Grunau
    orcid: 0000-0002-5428-9212
    corresponding: false
    email: rgrunau@mail.ubc.ca
    roles:
      - writing - review & editing
      - funding acquisition
    affiliations:
      - ref: 2
      - ref: 3
  - name: Alexander Mark Weber
    orcid: 0000-0001-7295-0775
    corresponding: true
    email: aweber@bcchr.ca
    roles:
      - project administration
      - supervision
      - validation
      - visualization
      - resources
      - methodology
      - formal analysis
      - funding acquisition
      - writing - review & editing
      - conceptualization
      - data curation
      - investigation
    affiliations:
      - ref: 2
      - ref: 3
    degrees:
      - PhD
      - MSc
affiliations:
  - id: 1
    name: Integrated Sciences, The University of British Columbia, Vancouver, BC, Canada
  - id: 2
    name: BC Children's Hospital Research Institute, The University of British Columbia, Vancouver, BC, Canada
  - id: 3
    name: Pediatrics, The University of British Columbia, Vancouver, BC, Canada
  - id: 4
    name: Physics and Astronomy, The University of British Columbia, Vancouver, BC, Canada

@@

notebook-links: false
number-sections: false
citation:
  container-title: Unpublished
# date: last-modified
bibliography: Gavin_Thesis_Ref.bib
csl: nature.csl

-   This study evaluated the use of QSM and its paramagnetic components to measure cerebral oxygenation in neonates.
-   By comparing susceptibility-derived oxygen saturation (SvO~2~) in the superior sagittal sinus (SSS) and central cerebral veins (CCV), it adds to the field of neonatal cerebral oxygenation measurement.
-   Decomposing QSM into paramagnetic components shows potential for improving SvO~2~ accuracy, particularly in the SSS, though variability remains a challenge.
-   The results suggest no significant oxygenation difference between the SSS and CCV, contrasting with previous findings, indicating a need for further research on neonatal venous oxygenation.

## Abstract:

**Background**: Quantitative susceptibility mapping (QSM), a magnetic resonance imaging (MRI) modality sensitive to deoxyhemoglobin, is a promising method for measuring cerebral oxygenation in human neonates. Paramagnetic sources, like deoxyhemoglobin, however, can be obscured by diamagnetic sources such as water and myelin. This study evaluated whether QSM images, or isolated paramagnetic components, are more accurate for measuring oxygenation of cerebral veins of preterm neonates, and explored oxygenation differences between the major cerebral veins.

@@

**Conclusion**: While paramagnetic component decomposition yielded SSS values closer to those found in the literature, it increased variability. No significant oxygenation differences were found between the SSS and CCV, contrasting with prior studies.

@@

@@

For the preceding reasons, non-invasive MRI-based techniques are actively being explored to assess regional and whole-brain blood oxygenation. While MRI-based methods have been developed for adults [@jainRapidMagneticResonance2011; @luQuantitativeEvaluationOxygenation2008; @xuNoninvasiveQuantificationWholebrain2009], their application in neonates is only beginning to be explored [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jainCerebralOxygenMetabolism2014; @jiangVesselspecificQuantificationNeonatal2019]. This delay is likely due to the unique challenges posed by neonates, including their smaller anatomies, distinct hemodynamic profiles, susceptibility to motion artifacts, and the difficulties associated with recruiting this population for research. These methods have almost all relied on T2 relaxation to estimate CSvO~2~ [@devisNoninvasiveMRIMeasurements2014; @liuQuantitativeAssessmentGlobal2014; @qiHemodynamicMetabolicAssessment2018; @jiangVesselspecificQuantificationNeonatal2019] with the exception of @jainCerebralOxygenMetabolism2014 which used susceptometry [@jainMRIEstimationGlobal2010]. One limitation of these T2 relaxation methods, however, is the fact that SvO~2~ is often measured using a single imaging slice, averaging values across several voxels, and only in the superior sagittal sinus (SSS). In the case of @jainCerebralOxygenMetabolism2014, they obtained regional and whole-brain data, but with thick slices (5mm), and still only estimated CSvO~2~ in the SSS. An alternative MRI method using quantitative susceptibility mapping (QSM) has been proposed, which can measure SvO~2~ regionally and across the whole-brain at high resolution (\< 1mm^3^ per voxel) [@weberQuantitativeSusceptibilityMapping2021]. However, this method left room for improvement, as it removed the SSS (averaging CSvO~2~ across the internal veins), and required an arbitrary threshold value of 0.15 ppm in order to acquire realistic results [@weberQuantitativeSusceptibilityMapping2021]. Furthermore, QSM tends to underestimates parametric components due to the inclusion of diamagnetic tissue, and vice versa, as the opposing magnetic susceptibilities effectively subtract from one another [@kimHSeparationImagingDiagnosis2023].

@@

Participant data comes from a previous study \*\*\* <!-- [@zhu-etal-cmro2] (submitted) -->. Participants consisted of preterm neonates born between 25- and 31-weeks gestational age (GA) who were admitted to the level III NICU at \*\*\* <!-- BC Women’s Hospital -->. Recruitment took place over a span of one year, from February 2021 to January 2022, facilitated by a dedicated research nurse. Parents of eligible infants were approached by the research nurse prior to discharge from the NICU to explain the study objectives and seek their consent for participation. Infants meeting the criteria for inclusion were scanned for the study if they had already been discharged from the NICU, were in stable condition, and had reached a term equivalent age of 37 to 44 weeks GA. However, certain exclusion criteria were applied to ensure the homogeneity and integrity of the study sample: infants were excluded if there was clinical evidence of a congenital malformation or syndrome, a TORCH infection, or ultrasound evidence of large parenchymal hemorrhagic infarction (\>2 cm, Grade 4 intraventricular hemorrhage).

@@

MR imaging was performed on a 3.0 Tesla General Electric Discovery MR750 scanner (scanner software version DV26.0_R03) equipped with a SREE Medical Systems (Cleveland, OH) single-channel neonatal head coil (@tbl-mri). The scans were conducted at the \*\*\* <!-- BC Children -->’s MRI Research Facility. Prior to the scanning procedure, subjects were carefully prepared by a research nurse. Swaddling and feeding were used to ensure the comfort and cooperation of the subjects during the scan. Importantly, no sedatives or invasive markers were utilized throughout the procedure. Subjects were placed within a specially designed SREE Medical Systems MRI compatible incubator, which facilitated both safety and motion minimization. Molded foam was strategically positioned around the head and body within the incubator to further restrict subject movement. To protect against potential hearing damage, ear plugs were employed during the scanning process. Additionally, a pulse oximeter was affixed to the subject’s foot to monitor arterial oxygen saturation and heart rate throughout the scan.

```{r}
#| label: tbl-mri
#| tbl-cap: "**Technical parameters for MR imaging pulse sequences**"
df <- data.frame(Scan = character(), T1w = character(), T2w = character(), pcASL = character(), SWI = character(), stringsAsFactors = FALSE)
df[1,] <- c("Sequence", "3D FSPGR", "3D CUBE", "Multi-shot 3D fast spin-echo", "3D spoiled GRE flow-compensated")
df[2,] <- c("Acquisition plane", "Coronal", "Sagittal", "Axial", "Axial")
df[3,] <- c("Phase-encoding direction", "Left-Right", "Posterior-Anterior", "Posterior-Anterior", "Left-Right")
df[4,] <- c("TR (ms)", "7.74", "2,300", "4,680", "30.9")
df[5,] <- c("TE (ms)", "2.97", "66.29", "10.55", "5 echoes; first echo: 5; echo spacing: 5.24")
df[6,] <- c("Flip angle", "12\U00B0", "90\U00B0", "111\U00B0", "20\U00B0")
df[7,] <- c("FOV (cm)", "20", "20", "24", "25")
df[8,] <- c("Acquisition matrix", "512 x 512", "256 x 256", "128 x 128", "256 x 256")
df[9,] <- c("In-plane resolution (mm)", "0.39 x 0.39", "0.78 x 0.78", "1.875 x 1.875", "0.977 x 0.977")
df[10,] <- c("Slice thickness (mm)", "1", "1", "4", "2, reconstructed to 1 with zero filling (ZIP2)")
df[11,] <- c("Number of slices", "126", "106", "50", "92")
df[12,] <- c("Additional parameters", "n/a", "n/a", "1,450 ms label period;\n 2,025 ms pulse label;\n 24 control-label pairs", "n/a")
df[13,] <- c("Scan duration", "4 min 39 s", "5 min 1 s", "5 min 26 s", "5 min 29 s")

@@

if (knitr::is_latex_output()) {
#colnames(df)[1] <- ""
colnames(df) <- c("", "\\textbf{T1w}", "\\textbf{T2w}", "\\textbf{pcASL}", "\\textbf{SWI}")
df[12,] <- linebreak(df[12,])
df |>
  kbl(format = "latex",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    align = "l",
    escape = FALSE) |>
    kable_styling(font_size = 8, position = "center", latex_options = c("hold_position", "scale_down"), full_width = FALSE) |>
    footnote(general_title = "", 
           footnote_as_chunk = TRUE,
           threeparttable = TRUE,
           general = footnotetext) |>
    column_spec(1, width="8em") |>
    column_spec(4, width="9em") |>
    column_spec(5, width="9em")
} else {
df |> 
    mutate(across(everything(), ~ str_replace_all(., "\n", "<br>"))) |> 
    gt(rowname_col = "Scan") |> tab_footnote(footnotetext) |>
    fmt_markdown(columns = everything()) |>
    cols_label(
      T1w = md("**T1w**"),
      T2w = md("**T2w**"),
      pcASL = md("**pcASL**"),
      SWI = md("**SWI**")
    ) |>
    tab_options(quarto.disable_processing = TRUE)
}
```

The MRI scan protocol comprised of the following sequences (plane of acquisition in parentheses): a T1-weighted scan (coronal), a T2-weighted scan (sagittal), a pseudo-continuous arterial spin labeling (ASL) scan [@alsopRecommendedImplementationArterial2015] (axial), a multi-echo susceptibility-weighted imaging scan [@denkSusceptibilityWeightedImaging2010] (axial), and a diffusion-weighted imaging (DWI) spin-echo echo planar imaging (EPI) sequence (axial). The DWI sequence was not used for the present study.

@@

The raw DICOM files acquired from the scanning procedure were converted to NIfTI (Neuroimaging Informatics Technology Initiative) format using Chris Rorden’s `dcmniix` tool [@liFirstStepNeuroimaging2016]. SWI magnitude data files were then used to create subject-specific brain masks that would not erode the SSS during QSM processing, an issue faced by our group in the past \*\*\* <!-- [@weberQuantitativeSusceptibilityMapping2021] -->. A step-by-step summary of the pipeline used is shown in @fig-graph.
